---
title: "Projection covid-19 cases in Ontario"
output: flexdashboard::flex_dashboard
runtime: shiny

---


```{r,echo=FALSE,warning=FALSE,message=FALSE,results='hide'}
library(DT)
library(fs)
library(gridExtra)
library(tidyverse)
library(dplyr)
library(data.table)
library(mgcv)
library(forecast)

library(directlabels)
library(ggthemes)
library(hexbin)
library(plotly)
library(quadprog)
#library(shinyWidgets)
#library(shinydashboard)

source("utils.R")
# #read raw github data from the working group github;

#on_reported <- fread("https://raw.githubusercontent.com/ishaberry/Covid19Canada/master/timeseries_prov/cases_timeseries_prov.csv") 


DownloadCOVIDdata_reported<-function(){
  download.file(
    url = "https://raw.githubusercontent.com/ishaberry/Covid19Canada/master/timeseries_prov/cases_timeseries_prov.csv",
    destfile = "data/OntarioReported.csv")
}


# update reported data
UpdateMyData_reported<-function(){
  T_refresh = 12 # hours 
  if(!dir_exists("data")){
    dir.create("data")
    DownloadCOVIDdata_reported()
  }
  else if((!file.exists("data/OntarioReported.csv") || as.double(Sys.time()- file_info("data/OntarioReported.csv")$change_time,units="hours") > T_refresh)){
    DownloadCOVIDdata_reported()
  }  
}



UpdateMyData_reported()


# Read confirmed data Ontario.data
DownloadCOVIDdata<-function(){
  download.file(
    url = "https://data.ontario.ca/dataset/f4112442-bdc8-45d2-be3c-12efae72fb27/resource/455fd63b-603d-4608-8216-7d8647f43350/download/conposcovidloc.csv",
    destfile = "data/OntarioConfirmed.csv")
}

# update confirmed data
UpdateMyData<-function(){
  T_refresh = 12 # hours 
  if(!dir_exists("data")){
    dir.create("data")
    DownloadCOVIDdata()
  }
  else if((!file.exists("data/OntarioConfirmed.csv") || as.double(Sys.time()- file_info("data/OntarioConfirmed.csv")$change_time,units="hours") > T_refresh)){
    DownloadCOVIDdata()
  }  
}
  
UpdateMyData() # updating the data in every 12 hours

```



Projection
=================================================
```{r}



# for confirmed cases
on_confirmed<-read_csv("data/OntarioConfirmed.csv")

#on_confirmed <- fread('https://data.ontario.ca/dataset/f4112442-bdc8-45d2-be3c-12efae72fb27/resource/455fd63b-603d-4608-8216-7d8647f43350/download/conposcovidloc.csv')
on_case_by_date<-on_confirmed%>%select(Date=Accurate_Episode_Date)%>%group_by(Date)%>%summarise(newcase=n())
on_data<-on_case_by_date%>%mutate(Date=as.Date(Date))%>%filter(Date >= as.Date("2020-02-15"))
on_data<-on_data%>%mutate(Day=as.numeric(as.Date(Date)-as.Date("2020-02-14")),cum_case=cumsum(newcase))
#case<-ts(on_data[,4],start=1,end=60,frequency=1) # creating time-series data



# for reported cases
reported<-read_csv("data/OntarioReported.csv")
on_reported<-reported%>%filter(province=="Ontario")%>%select(Date=date_report,newcase=cases)
on_data_reported<-on_reported[-(1:27),]
on_data_reported<-on_data_reported%>%mutate(Day=1:nrow(on_data_reported),cum_case=cumsum(newcase))
#fit <- tslm(log(case) ~ trend)
#fc<-forecast(fit, h=5)
#plot(fc)
#summary(fit)

#plot(forecast(fit, fan = TRUE))


b2<-gam(newcase ~ s(Day,bs="tp",k=20),family=nb(),data = on_data,method = "REML")


b2_rep<-gam(newcase ~ s(Day,bs="tp",k=20),family=nb(),data = on_data_reported,method = "REML")


# agrument takes upto next 7 days
Projection<-function(Day_upto)
{
  if(Day_upto %in% 1:5)
  {
    Day<-seq(on_data$Day[nrow(on_data)]+1,on_data$Day[nrow(on_data)]+Day_upto,by=1)
    newdata_c<-as.data.frame(Day)
    coPred = predict(b2, newdata_c, se.fit = TRUE)
    newdata_c <- within(newdata_c, {
      Predicted_cases_conf <- exp(coPred$fit)
      LL_c <- exp(coPred$fit - 1.96 * coPred$se.fit)
      UL_c <- exp(coPred$fit + 1.96 * coPred$se.fit)
    })
    
    Day<-seq(on_data_reported$Day[nrow(on_data_reported)]+1,on_data_reported$Day[nrow(on_data_reported)]+Day_upto,by=1)
    newdata_r<-as.data.frame(Day)
    coPred_rep = predict(b2_rep, newdata_r, se.fit = TRUE)
    newdata_r <- within(newdata_r, {
      Predicted_cases_rep <- exp(coPred_rep$fit)
      LL_r <- exp(coPred_rep$fit - 1.96 * coPred_rep$se.fit)
      UL_r <- exp(coPred_rep$fit + 1.96 * coPred_rep$se.fit)
    })
    
    
    
    
    
    p<-ggplot(on_data, aes(x = Day, y = newcase)) + geom_point(size=0.5) +geom_line(colour='blue')+
      theme_bw()+ geom_smooth(aes(x=Day, y=Predicted_cases_conf, ymax=UL_c, ymin=LL_c), colour='red', data=newdata_c, stat='identity')+
      geom_point(aes(x=Day,y=Predicted_cases_conf),data=newdata_c)+geom_smooth()+
      scale_x_continuous(breaks=c(0,20,40,60,80), labels=c("Feb 15", "Mar 05", "Mar 25", "Apr 14", "May 05"))+
      annotate("text",x=newdata_c$Day[nrow(newdata_c)]+2,y=ceiling(newdata_c$Predicted_cases_conf[nrow(newdata_c)]+60),label = ceiling(newdata_c$Predicted_cases_conf[nrow(newdata_c)]))+
      theme(axis.title=element_text(size=14,face="bold"))+ylab("Confirmed covid-19 cases in Ontario")+xlab("Date")
    
    
    q<-ggplot(on_data_reported, aes(x = Day, y = newcase)) + geom_point(size=0.5) +geom_line(colour='blue')+
      theme_bw()+ geom_smooth(aes(x=Day, y=Predicted_cases_rep, ymax=UL_r, ymin=LL_r), colour='red', data=newdata_r, stat='identity')+
      geom_point(aes(x=Day,y=Predicted_cases_rep),data=newdata_r)+geom_smooth()+
      scale_x_continuous(breaks=c(0,20,40,60), labels=c("Feb 20", "Mar 10", "Mar 30", "Apr 20"))+
      annotate("text",x=newdata_r$Day[nrow(newdata_r)]+2,y=ceiling(newdata_r$Predicted_cases_rep[nrow(newdata_r)]+60),label = ceiling(newdata_r$Predicted_cases_rep[nrow(newdata_r)]))+
      theme(axis.title=element_text(size=14,face="bold"))+ylab("Reported covid-19 cases in Ontario")+xlab("Date")
    final_plot<-grid.arrange(p, q, nrow = 1)
    print(final_plot)
  }
  else{print("projection is not available beyond 5 days")}
}




```


Inputs {.sidebar}
-------------------------------------


```{r}
sliderInput("Day_upto", 
                  div(HTML("Projection up to next 5 days")), 
                  min   = 1, 
                  max   = 5, 
                  value = 3, 
                  step  = 1)
```

Row
--------------------------------------------------

```{r}
      plotOutput("distPlot", height = "500px")
      
output$distPlot <- renderPlot({
    Projection(input$Day_upto)
  })

```


The actual cases are presented using blue lines and projections are in red. Last projected cases are printed in the plot. Note that it takes about a week for Ontario officials to confirm test results. Thus, the last week's confirmed counts are different compared to other sources. The projections are estimated using a simple Generalized Additive model with only one covariate, wihch is Time. The performance of the model can be improved by adding important convariates into it.  

About
=================================================
This is a simple app to facilitate Public Health Ontario to estimate staff needed for contact tracing and resource allocation. 

- **Template and code**
  - The Template has been forked from Tinu Schneider's [github](https://github.com/tinu-schneider/Flatten_the_Curve) and the code are avaialble [here](https://github.com/sahasud1/covid19).
  
- **Data**
  - The data is available for download [here](https://data.ontario.ca/dataset/f4112442-bdc8-45d2-be3c-12efae72fb27/resource/455fd63b-603d-4608-8216-7d8647f43350/download/conposcovidloc.csv).

- **Data Source**
  - Data source is [here](https://data.ontario.ca/dataset/).
  
- **Feedbacks**
  - Please let me know if you have any suggestions on how to improve the projection. You to reach me with at [sudip.saha@mail.utoronto.ca](mailto:sudip.saha@mail.utoronto.ca). More about [me](https://sahasud1.github.io/).


